---
title: "Higher Temperatures lead to Masting Years"
author: "Justin Beslity"
date: "August 3, 2019"
output: 
  ioslides_presentation:
    widescreen: true
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
```

##Packages
```{r echo=F}
library(tidyverse)
library(zoo)
library(lubridate)
library(anchors)
```


```{r echo=F}
weather <- read.csv("../data_raw/atmos.csv",  header = T)

NCDC <- read.csv("../data_raw/NCDC_data.csv", header = T)
Newcomb <- NCDC[which(NCDC$NAME == "NEWCOMB, NY US"),]

weather$Tavg <- (weather$TMAX+weather$TMIN)/2
Newcomb$Tavg <- (Newcomb$TMAX+Newcomb$TMIN)/2
```


##Break Weather into seasons
```{r echo= F}

Newcomb$quarter <- as.yearqtr(as.yearmon(Newcomb$DATE, "%Y-%m-%d") + 1/12)
Newcomb$seasons <- factor(format(Newcomb$quarter, "%q"), levels = 1:4, 
                labels = c("winter", "spring", "summer", "fall"))

```


##Aggregate average temperature for the growing season
```{r echo =F}
w_grow <- Newcomb[which(Newcomb$seasons == "summer"),]
w_grow$YEAR <- year(w_grow$DATE)

w_grow_y_avg <- aggregate(w_grow$Tavg ~ w_grow$YEAR, FUN = "mean")
w_grow_y_max <- aggregate(w_grow$TMAX ~ w_grow$YEAR, FUN = "mean")
w_grow_y_pre <- aggregate(w_grow$PRCP ~ w_grow$YEAR, FUN = "sum")
```



##dT Methods
```{r}
for(i in seq(1:nrow(w_grow_y_avg))){
w_grow_y_avg[i+2,3] <- w_grow_y_avg[i+1,2] - w_grow_y_avg[i,2]
}

for(i in seq(1:nrow(w_grow_y_max))){
w_grow_y_max[i+2,3] <- w_grow_y_max[i+1,2] - w_grow_y_max[i,2]
}

colnames(w_grow_y_avg) <- c("year", "MeanAvgTemp", "avg_dT_C")
colnames(w_grow_y_max) <- c("year", "MeanMaxTemp", "max_dT_C")

```


#Trim weather data to match seed data
```{r echo=F}
w_grow_y_avg <- w_grow_y_avg[-c(1:4),]
w_grow_y_avg <- w_grow_y_avg[-c(32:34),]

w_grow_y_max <- w_grow_y_max[-c(1:4),]
w_grow_y_max <- w_grow_y_max[-c(32:34),]

w_grow_y_pre_cur <- w_grow_y_pre[-c(1:29),]
w_grow_y_pre_cur <- w_grow_y_pre_cur[-c(32),]

#Previous Year Precip
w_grow_y_pre_prev <- w_grow_y_pre[-c(1:28),]
w_grow_y_pre_prev <- w_grow_y_pre_prev[-c(32:33),]

colnames(w_grow_y_pre_cur) <- c("Year", "Precip_in")
colnames(w_grow_y_pre_prev) <- c("Year", "Precip_prev_in")


write.csv(w_grow_y_avg, "../data_processed/2dT_avgDailyTemps.csv")
write.csv( w_grow_y_max, "../data_processed/2dT_MaxDailyTemps.csv")
        

```


##Bird Temp Data (Not important right now)
```{r, echo=F}
w_t <- weather[which(weather$MONTH == 4 & 5 & 6), ]

w_t_avg <- aggregate(w_t$Tavg ~ w_t$YEAR, FUN = "mean")
colnames(w_t_avg) <- c("year", "Tavg_C")

w_t_max <- aggregate(w_t$TMAX ~ w_t$YEAR, FUN = "mean")
colnames(w_t_max) <- c("year", "Tmax_C")

write.csv(w_t_avg, "bird_temp_AMJ.csv")
write.csv(w_t_max, "bird_temp_maxDaily_AMJ.csv")


```


##Aggregate Seed Data by tree type
```{r eval = T, echo=F}
seed <- read.csv("../data_raw/Seed_data.csv", header = TRUE)

seed_SUMA <- aggregate(seed$SUMA_TOTAL ~ seed$YEAR, FUN = "sum")
seed_AMBE <- aggregate(seed$AMBE_TOTAL ~ seed$YEAR, FUN = "sum")

ltm_suma <- mean(seed_SUMA$`seed$SUMA_TOTAL`)
ltm_ambe <- mean(seed_AMBE$`seed$AMBE_TOTAL`)

sd_suma <- sd(seed_SUMA$`seed$SUMA_TOTAL`)
sd_ambe <- sd(seed_AMBE$`seed$AMBE_TOTAL`)

seed_SUMA$dev <- (seed_SUMA$`seed$SUMA_TOTAL` - ltm_suma) / sd_suma
seed_AMBE$dev <- (seed_AMBE$`seed$AMBE_TOTAL` - ltm_ambe) / sd_ambe

thresh_suma <- -mean(seed_SUMA$dev[which(seed_SUMA$dev < 0)])
thresh_ambe <- -mean(seed_AMBE$dev[which(seed_AMBE$dev < 0)])

colnames(seed_SUMA) <- c("YEAR", "COUNT", "DEV")
colnames(seed_AMBE) <- c("YEAR", "COUNT", "DEV")
```


##Determining Sugar Maple Mast years
```{r eval=T, echo=F}
plot(seed_SUMA$YEAR, seed_SUMA$DEV, ylab = "", xlab = "")
abline(h = thresh_suma)
mtext("Annual Standardized Deviation", side = 2, line = 2.5)
mtext("Year", side = 1, line = 2.5)
mtext("Sugar Maple", side = 3, line = 1, cex = 1.5)

```


##Determining American Beech Mast years
```{r eval=T, echo=F}
plot(seed_AMBE$YEAR, seed_AMBE$DEV, ylab = "", xlab = "")
abline(h = thresh_ambe)
mtext("Annual Standardized Deviation", side = 2, line = 2.5)
mtext("Year", side = 1, line = 2.5)
mtext("American Beech", side = 3, line = 1, cex = 1.5)

```

##Combining weather data and removing Post Mast years
```{r eval = T, echo=F}

seed_AMBE <- cbind(seed_AMBE, w_grow_y_avg, w_grow_y_max, w_grow_y_pre_cur, w_grow_y_pre_prev)
seed_AMBE <- seed_AMBE[,-c(4,7)]

seed_SUMA <- cbind(seed_SUMA, w_grow_y_avg, w_grow_y_max, w_grow_y_pre_cur, w_grow_y_pre_prev)
seed_SUMA <- seed_SUMA[,-c(4,7)]

suma_cor <- cor(seed_SUMA$COUNT, seed_SUMA$avg_dT_C)
AMBE_cor <- cor(seed_AMBE$COUNT, seed_AMBE$avg_dT_C)

mast_y_suma <- seed_SUMA$YEAR[which(seed_SUMA$DEV > thresh_suma)]
post_mast_suma <- mast_y_suma+1
seed_SUMA_trim <- filter(seed_SUMA, !(YEAR %in% post_mast_suma))

mast_y_ambe <- seed_AMBE$YEAR[which(seed_AMBE$DEV > thresh_ambe)]
post_mast_ambe <- mast_y_ambe+1
seed_AMBE_trim <- filter(seed_AMBE, !(YEAR %in% post_mast_ambe))


```

## 1a) Sugar Maple seed count w/ post mast years mean average daily temperature
```{r eval=T, echo=F}
SUMA_a_lm <- lm(seed_SUMA$COUNT ~ seed_SUMA$avg_dT_C)
summary(SUMA_a_lm)
plot(seed_SUMA$avg_dT_C, seed_SUMA$COUNT, ylab = "", xlab = "")
abline(a = 194.64, b = 62.02)
legend(legend = "adj r2 = 0.32", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("Sugar Maple", side = 3, line = 1, cex = 1.5)

```


## 1b) American Beech seed count w/ post mast years mean average daily temperaure
```{r eval = T, echo=F}
AMBE_a_lm <- lm(seed_AMBE$COUNT ~ seed_AMBE$avg_dT_C)
summary(AMBE_a_lm)
plot(seed_AMBE$avg_dT_C, seed_AMBE$COUNT, ylab = "", xlab = "")
abline(a = 46.22, b = 12.64)
legend(legend = "adj r2 = 0.34", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("American Beech", side = 3, line = 1, cex = 1.5)
```


## 2a) Sugar Maple seed count w/ post mast years mean maximum daily temperature
```{r eval=T, echo=F}
SUMA_m_lm <- lm(seed_SUMA$COUNT ~ seed_SUMA$max_dT_C)
summary(SUMA_m_lm)
plot(seed_SUMA$avg_dT_C, seed_SUMA$COUNT, ylab = "", xlab = "")
abline(a = 194.64, b = 62.02)
legend(legend = "adj r2 = 0.39", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("Sugar Maple", side = 3, line = 1, cex = 1.5)

```


## 2b) American Beech seed count w/ post mast years mean maximum daily temperaure
```{r eval = T, echo=F}
AMBE_m_lm <- lm(seed_AMBE$COUNT ~ seed_AMBE$max_dT_C)
summary(AMBE_m_lm)
plot(seed_AMBE$avg_dT_C, seed_AMBE$COUNT, ylab = "", xlab = "")
abline(a = 46.22, b = 12.64)
legend(legend = "adj r2 = 0.41", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("American Beech", side = 3, line = 1, cex = 1.5)
```



## 3a) Sugar Maple seed count w/o post mast years mean average daily temperatures
```{r}
SUMA_t_a_lm <- lm(seed_SUMA_trim$COUNT ~ seed_SUMA_trim$avg_dT_C)
summary(SUMA_t_a_lm)
plot(seed_SUMA_trim$avg_dT_C, seed_SUMA_trim$COUNT, ylab = "", xlab = "")
abline(a = 169.21, b = 91.24)
legend(legend = "adj r2 = 0.3", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("Sugar Maple", side = 3, line = 1, cex = 1.5)

```


# 3b) American Beech seed count w/o post mast years mean average daily temperatures
```{r eval = T, echo=F}
AMBE_t_a_lm <- lm(seed_AMBE_trim$COUNT ~ seed_AMBE_trim$avg_dT_C)
summary(AMBE_t_a_lm)
plot(seed_AMBE_trim$avg_dT_C, seed_AMBE_trim$COUNT, ylab = "", xlab = "")
abline(a = 46.22, b = 12.64)
legend(legend = "adj r2 = 0.34", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("American Beech", side = 3, line = 1, cex = 1.5)
```


## 4a) Sugar Maple seed count w/o post mast years mean maximum daily temperatures
```{r}
SUMA_t_m_lm <- lm(seed_SUMA_trim$COUNT ~ seed_SUMA_trim$max_dT_C)
summary(SUMA_t_m_lm)
plot(seed_SUMA_trim$max_dT_C, seed_SUMA_trim$COUNT, ylab = "", xlab = "")
abline(a = 169.21, b = 91.24)
legend(legend = "adj r2 = 0.47", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("Sugar Maple", side = 3, line = 1, cex = 1.5)

```


# 4b) American Beech seed count w/o post mast years mean maximum daily temperatures
```{r eval = T, echo=F}
AMBE_t_m_lm <- lm(seed_AMBE_trim$COUNT ~ seed_AMBE_trim$max_dT_C)
summary(AMBE_t_m_lm)
plot(seed_AMBE_trim$max_dT_C, seed_AMBE_trim$COUNT, ylab = "", xlab = "")
abline(a = 46.22, b = 12.64)
legend(legend = "adj r2 = 0.42", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("American Beech", side = 3, line = 1, cex = 1.5)
```


## 5a) Sugar Maple seed count w/o post mast years mean maximum daily temperatures + PRECIP of current year
```{r}
SUMA_t_m_lm <- lm(seed_SUMA_trim$COUNT ~ seed_SUMA_trim$max_dT_C + seed_SUMA_trim$Precip_in)
summary(SUMA_t_m_lm)
plot(seed_SUMA_trim$max_dT_C, seed_SUMA_trim$COUNT, ylab = "", xlab = "")
abline(a = 169.21, b = 91.24)
legend(legend = "adj r2 = 0.45", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("Sugar Maple", side = 3, line = 1, cex = 1.5)

```


# 5b) American Beech seed count w/o post mast years mean maximum daily temperatures + PRECIP of current year
```{r eval = T, echo=F}
AMBE_t_m_lm <- lm(seed_AMBE_trim$COUNT ~ seed_AMBE_trim$max_dT_C + seed_AMBE_trim$Precip_in)
summary(AMBE_t_m_lm)
plot(seed_AMBE_trim$max_dT_C, seed_AMBE_trim$COUNT, ylab = "", xlab = "")
legend(legend = "adj r2 = 0.52", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("American Beech", side = 3, line = 1, cex = 1.5)

res_AMBE_t_m_lm <- residuals(AMBE_t_m_lm)
qqnorm(res_AMBE_t_m_lm)
qqline(res_AMBE_t_m_lm)
```

## 6a) Sugar Maple seed count w/o post mast years mean maximum daily temperatures + PRECIP of previous year
```{r}
SUMA_t_m_lm <- lm(seed_SUMA_trim$COUNT ~ seed_SUMA_trim$max_dT_C + seed_SUMA_trim$Precip_prev_in)
summary(SUMA_t_m_lm)
plot(seed_SUMA_trim$max_dT_C, seed_SUMA_trim$COUNT, ylab = "", xlab = "")
abline(a = 169.21, b = 91.24)
legend(legend = "adj r2 = 0.48", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("Sugar Maple", side = 3, line = 1, cex = 1.5)

res_SUMA_t_m_lm <- residuals(SUMA_t_m_lm)
qqnorm(res_SUMA_t_m_lm)
qqline(res_SUMA_t_m_lm)

```


# 6b) American Beech seed count w/o post mast years mean maximum daily temperatures + PRECIP of previous year
```{r eval = T, echo=F}
AMBE_t_m_lm <- lm(seed_AMBE_trim$COUNT ~ seed_AMBE_trim$max_dT_C + seed_AMBE_trim$Precip_prev_in)
summary(AMBE_t_m_lm)
plot(seed_AMBE_trim$max_dT_C, seed_AMBE_trim$COUNT, ylab = "", xlab = "")
abline(a = 46.22, b = 12.64)
legend(legend = "adj r2 = 0.40", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("American Beech", side = 3, line = 1, cex = 1.5)
```

# 7) ln(American Beech seed count) w/o post mast years mean maximum daily temperatures + PRECIP of current year
```{r eval = T, echo=F}
lnAMBE_t_m_lm <- lm(log(seed_AMBE_trim$COUNT) ~ seed_AMBE_trim$max_dT_C + seed_AMBE_trim$Precip_in)
summary(lnAMBE_t_m_lm)
plot(seed_AMBE_trim$max_dT_C, log(seed_AMBE_trim$COUNT), ylab = "", xlab = "")
legend(legend = "adj r2 = 0.62", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("American Beech", side = 3, line = 1, cex = 1.5)

res_lnAMBE_t_m_lm <- residuals(lnAMBE_t_m_lm)
qqnorm(res_lnAMBE_t_m_lm)
qqline(res_lnAMBE_t_m_lm)

```

## 8) ln(Sugar Maple seed count) w/o post mast years mean maximum daily temperatures + PRECIP of previous year
```{r}
#Gotta change 0's to 1
lnSUMA <- seed_SUMA_trim
lnSUMA[13,2] <- 1
lnSUMA[24,2] <- 1

lnSUMA_t_m_lm <- lm(log(lnSUMA$COUNT) ~ lnSUMA$max_dT_C + lnSUMA$Precip_prev_in)
summary(lnSUMA_t_m_lm)
plot(seed_SUMA_trim$max_dT_C, log(seed_SUMA_trim$COUNT), ylab = "", xlab = "")
legend(legend = "adj r2 = 0.37", "topleft", cex = 1.2, bty = "n")
mtext("Differential Temperature (C)", side = 1, line = 2.5 )
mtext("Seed Count", side = 2, line = 2.5 )
mtext("Sugar Maple", side = 3, line = 1, cex = 1.5)

res_lnSUMA_t_m_lm <- residuals(lnSUMA_t_m_lm)
qqnorm(res_lnSUMA_t_m_lm)
qqline(res_lnSUMA_t_m_lm)

```