---
title: "EDS_Birds"
author: "Victoria_Field"
date: "8/2/2019"
output:
  ioslides_presentation:
    incremental: true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```

# Migratory Birds (1984-2008)

## Project Background

-bird counts (5-6) were conducted in early June annually on 4 unlimited- radius plots per site on 4 sites

-intent was to compare forest management types between seedling-sapling hardwood (Maple Sale), sapling-pole hardwood (Sucker Brook), large pole-small sawtimber mixed (Adjidaumo) (1983-86), red spruce (Picea rubens) pole (Hare Area), and old-growth hardwood (Natural Area) stands

## Read in Birds Data 

```{r, echo=FALSE}
Birds<-read.csv("../data_raw/ALTEMP2_Songbirds_1952_2017.csv", header=TRUE)
```

```{r, echo=TRUE}
str(Birds)
```

## Read in Species Names

```{r}
Species<-read.csv("bird_species_diet.csv", header=TRUE, na.strings=c("", " "), strip.white=TRUE)
colnames(Species)[colnames(Species)=="four_code"] <- "species"
```

## Add full name and diet guild to species

```{r}
#diet reflects primary feeding guild in summertime (taken from Birds of North America; Cornell Lab of Ornithology)  
Birds<-merge(Birds, Species, by="species", all.x = TRUE)
```

# Bird data exploration 

## Summary

```{r}
colnames(Birds)[3]<-"YEAR"
summary(Birds)
```

## Descriptions

```{r, echo=FALSE, eval=TRUE, warning=FALSE}
library(Hmisc)
Birds.desc<-describe(Birds)
```


```{r, echo=FALSE, eval=TRUE}
print(Birds.desc)
```

# Data summaries and plots

## Aggregating the data

```{r}
B_year_area<-aggregate(count~YEAR+area, Birds, FUN=sum)
B_year_diet<-aggregate(count~YEAR+diet, Birds, FUN=sum)
B_year<-aggregate(count~YEAR, Birds, FUN=sum)
```

## Counts by area

```{r, echo=FALSE, eval=TRUE}
ggplot(B_year_area, aes(x=YEAR, y=count))+
  geom_point(aes(x=YEAR, y=count, color=area))+
  labs(title="Total counts across all species for each area")+
  xlab("Year")+
  theme(axis.text.x = element_text(angle=60, hjust=1))
  ylab("Total Bird Count")
```

## Counts by diet

```{r, echo=FALSE, eval=TRUE}
ggplot(B_year_diet, aes(x=YEAR, y=count))+
  geom_point(aes(x=YEAR, y=count, color=diet))+
  labs(title="Total counts for each diet guild")+ 
  xlab("Year")+
  theme(axis.text.x = element_text(angle=60, hjust=1))
  ylab("Total Bird Count")
```

## Counts by year

```{r, echo=FALSE, eval=TRUE}
ggplot(B_year, aes(x=YEAR, y=count))+
  geom_point(aes(x=YEAR, y=count))+
  labs(title="Total counts for each year")+
  xlab("Year")+
  theme(axis.text.x = element_text(angle=60, hjust=1))
  ylab("Total Bird Count")
```

# Temperature Data 

## Read in 

```{r}
Temp<-read.csv("../data_raw/atmos.csv")
Temp$TMEAN<-(Temp$TMAX+Temp$TMIN)/2
Temp$TDIFF<-(Temp$TMAX-Temp$TMIN)
```

```{r}
temp_month<-aggregate(TMEAN~YEAR+MONTH, FUN=mean, data=Temp)
temp_year<-aggregate(TMEAN~YEAR, FUN=mean, data=Temp)
```

```{r}
temp_year_diff<-aggregate(TDIFF~YEAR, FUN=mean, data=Temp)
```

## Merge birds and temperature 

```{r}
BirdTemp<-merge(temp_year, B_year, by="YEAR", all = TRUE)
BirdTemp<-merge(BirdTemp, temp_year_diff, by="YEAR", all =TRUE)
```

## Plot birds and temperature 

```{r}
par(mfrow=c(2,1))
plot(TDIFF~YEAR, BirdTemp)
plot(count~YEAR, BirdTemp)
```

```{r}
birdtemp.lm<-lm(count~TMEAN, BirdTemp)
plot(count~TMEAN, BirdTemp, xlim=c(38.7,44))
regLine(birdtemp.lm)
summary(birdtemp.lm)
```

```{r}
birdtempdiff.lm<-lm(count~TDIFF, BirdTemp)
plot(count~TDIFF, BirdTemp)
regLine(birdtempdiff.lm)
summary(birdtempdiff.lm)
```

# Seed Data

```{r}
Seed<-read.csv("../data_raw/Seed_data.csv", header=TRUE)
```

## Create new dataframe grouped by year and season where the sum of each column is computed.

```{r}
library(dplyr)
seed_year<-Seed[, -13] %>%
    group_by(YEAR, SEASON) %>% 
    summarise_each(funs(sum))
```

## Create new column with total seed counts and total seed biomass 

```{r}
seed_year$seed_count<-(seed_year$YEBI_TOTAL+seed_year$SUMA_TOTAL+seed_year$REMA_TOTAL+seed_year$AMBE_TOTAL+seed_year$RESP_TOTAL+seed_year$EAHE_TOTAL+seed_year$NWCE_TOTAL+seed_year$BAFI_TOTAL+seed_year$PABI_TOTAL)

seed_year$seed_biomass<-(seed_year$yebi_biomass+seed_year$suma_biomass+seed_year$rema_biomass+seed_year$ambe_biomass+seed_year$resp_biomass+seed_year$eahe_biomass+seed_year$nwce_biomass+seed_year$bafi_biomass+seed_year$pabi_biomass)
```

## Extract species of interest only (SUMA and AMBE)

```{r}
SppSeed<-seed_year[,c(1,2,5,7,14,16)]
```

## Plot 

## Assemblage Composition 

```{r}
comp<-aggregate(count~diet+YEAR+area, Birds, FUN=sum)
comp<-merge(comp, B_year_area, by=c("YEAR", "area"), all.x=TRUE)
comp$prop<-(comp$count.x/comp$count.y)
```

```{r}
ggplot(comp, aes(x=YEAR, y=count.x, group=diet))+
  geom_col(aes(color=diet, fill=diet))+
  labs(title="Assemblage Composition by Feeding Guild")+
  xlab("Year")+
  theme(axis.text.x = element_text(angle=60, hjust=1))+
  ylab("Total Bird Count")+
  facet_wrap(~area)
```